### 2025-04-09 manuscript/manuscript.Rmd
references

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Abigail Edwards"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "abigail.edwards@uq.edu.au"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Mikko Arvas"
    affiliation   : "2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "The University of Queensland, Australia"
  - id            : "2"
    institution   : "Finnish Red Cross Blood Service, Finland"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib", "r-references.bib"]
#create with zotero using "Better Bibtex" https://retorque.re/zotero-better-bibtex/

csl               : vox-sanguinis.csl
#https://www.zotero.org/styles?q=vox

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
knit              : worcs::cite_all
---

```{r setup, include = FALSE}
library("papaja")
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
# load_data()
r_refs("r-references.bib")
```

This manuscript uses the Workflow for Open Reproducible Code in Science [WORCS version 0.1.1, @lissaWORCSWorkflowOpen2021] to ensure reproducibility and transparency. All code <!--and data--> are available at <!--insert repository URL-->.

This is an example of a non-essential citation [@lissaWORCSWorkflowOpen2021]. If you change the rendering function to `worcs::cite_essential`, it will be removed.

In this project we aim to replicate the analysis carried out with Australian blood donors [@edwardsPsychologicalOwnershipIdentity2023] with Finnish blood donors

<!--The function below inserts a notification if the manuscript is knit using synthetic data. Make sure to insert it after load_data().-->
`r notify_synthetic()`

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

<!--https://tex.stackexchange.com/questions/702678/generating-citation-keys-with-zotero-and-better-bibtex-plugin  Zotero library -> right click -> Better BibTex -> Refresh BibTex Key. Do this for each entry. Then export the collection/library as references.bib file to /manuscript folder for use with this file -->

::: {#refs custom-style="Bibliography"}
:::