### 2025-06-10 manuscript_code.Rmd
multigroup analysis

---
title: 'Appendix 2: Manuscript and Exploratory Code'
author: "Abigail Edwards, Lauri Hietajärvi and Mikko Arvas"
date: "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time())`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(worcs)
library(tidyverse)
library(gtsummary)
library(likert)
library(here)
library(corrplot)
library(ggstats)
library(lavaan)
library(lavaanPlot)
library(kableExtra)
library(naniar)

library(gridExtra)
#library(semPlot) # update plots to semPlot later
#library(blavaan) # could not get it to work with this model and synthetic data, try as sensitivity analysis later
#library(tidySEM) # https://github.com/yrosseel/lavaan/issues/359 lavaan does not work if tidySEM is loaded!
# renv::status() to check 
# renv::snapshot() to add used and install to renv

# guide on how to setup worcs to reproduce this analysis https://cjvanlissa.github.io/worcs/articles/reproduce.html
```

# Data

```{r}
load_data()

# if (exists("data_simulated")) {
#   data <- data_simulated
#   rm(data_simulated)
# }
```

```{r}
# sometimes I get strange errors about C stack which can be solved with this
data <- as.data.frame(data)
fitmodels <- TRUE # TRUE for fitting models
#fitmodels <- FALSE # TRUE for fitting models
if (!fitmodels) {
  print("Loading previously fitted models")
  load(file=here("prereg_models.rdata"))
  fit.po02ss <- models$fit.po02ss
  fit.po03ss <- models$fit.po03ss
  fit.po04ss <- models$fit.po04ss
  fit.po05ss <- models$fit.po05ss
  fit.po02 <- models$fit.po02
  fit.po03 <- models$fit.po03
  fit.po04 <- models$fit.po04
  fit.po05 <- models$fit.po05
} else{
  models <- list()
}

dir.create(here("figures"))

```

```{r}
summary(data)
```

# Plot

##data cleaning

```{r eval = FALSE}
data %>% filter(bd04 < 0)
data %>% filter(bd03 > 12)

data <- data %>% 
  mutate(
    bd04 = ifelse(bd04 < 0, NA, bd04)/7,
    bd03 = ifelse(bd03 > 12, NA, bd03)
)

#afterdata changing negative blood donation history to missing data for bd04 from 3 participants
#afterdata changing values over 12 (maximum number of whole blood donations for men within 2 years) for bd03 to missing.
#afterdata rescaling bd04 to number of weeks since last donation
```

## Demographics Table 1

```{r eval = FALSE}

table1 <-
  tbl_summary(
    data = data %>% 
      rename(
      Gender =  de01,
      Age = de02,
      DonationEligibility = bd01,
      DonationCount = bd02,
      DonationCountLast2years = bd03,
      DaysFromLastDonation = bd04 
    ),
    include = c(
      Age,
      DonationEligibility,
      DaysFromLastDonation,
      DonationCount,
      DonationCountLast2years
      ),
    by =  Gender,
     statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", all_categorical() ~
    "{n} ({p}%)")

  ) %>%
  add_n()  # add column with total number of non-missing observations
as_kable_extra(table1, format = "latex")
```

## Age by sex

```{r eval=FALSE}

p <- ggplot(data %>% filter(de01 == 'man' | de01 == "woman"))
p <- p +  geom_histogram(aes(x=de02),binwidth = 5)
p <- p + facet_wrap(. ~ de01)
p <- p + xlab("Age in years")
p

```

## Psychological ownership

```{r eval=FALSE}
gglikert(
  data, 
  include =   starts_with("po") 
)
```

## Self-identity

```{r eval=FALSE}
gglikert(
  data, 
  include =   starts_with("si") 
)
```

## Intention

```{r eval=FALSE}

#how would we seperate out ie01; as it shouldn't sit with it01

gglikert(
  data,
  include = c("it01","it02","it03","ie01")
)
```

## Antecedents of psychological ownership - Control

```{r eval=FALSE}
gglikert(
  data, 
  include =   starts_with("ac") 
)
```

## Antecedents of psychological ownership - Intimate knowledge

```{r eval=FALSE}
gglikert(
  data, 
  include =   starts_with("ai") 
)
```

## Antecedents of psychological ownership - Self-Investment

```{r eval=FALSE}
gglikert(
  data, 
  include =   starts_with("iv") 
)
```

# Exclusions

```{r}

#removing non-eligible participants (must be: over 18, have donated blood in past 2 years, believe currently eligible to donate blood)

data %>%
  filter(!(de01 == 'man' | de01 == "woman"))  %>% nrow()
cat("are removed because of sex\n")

data %>%
  filter(bd01 == 'no') %>% nrow()
cat("are removed because of non-eligibility\n")

data %>%
  filter(bd02 < 1) %>% nrow()
cat("are removed because of non-donor\n")

data %>%
  filter(bd03 < 1) %>% nrow()
cat("are removed because of 2 year donation\n")

data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "woman") %>% 
  filter(bd01 != 'no') %>% 
  filter (bd02 >= 1) %>%
  filter (bd03 >= 1)%>% nrow()
cat("in total after exclusions.\n")

#afterdata adding additional exclusion criteria in line with preregistered exclusion for bd02 & bd03  
```

# Model

## Format data for modelling

```{r}
isordered <- data %>%  summarise(across(everything(), ~ is.ordered(.x))) %>%  as.logical() %>% which()
isfactor <- data %>%  summarise(across(everything(), ~ is.factor(.x))) %>%  as.logical() %>% which()

data <- data %>% 
  filter(bd02 >= 1)

data <- data %>% 
  filter(bd03 >= 1)

#standarsing behaviour variables
data <- data %>%
mutate(across(c(bd02, bd03, bd04), ~scale(.) %>% as.numeric()))

#lavaan does not understand empty factor levels
data.fit <- data %>%
  #lavaan cannot deal with unordered factor with more then 2 levels
  filter(de01 == 'man' | de01 == "woman") %>% # EXCLUSION
  filter(bd01 != 'no') %>% # EXCLUSION
  mutate(
    across(isfactor, ~ fct_drop(.x)
    ) %>% 
      mutate(
        bd04 = as.numeric(bd04), # lavaan cannot handle dates
        )
  ) %>% 
  #Original paper used sum scores instead of FA. 
  mutate(
    ss_psycown = 
      across(starts_with("po"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_selfid = 
      across(starts_with("si"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    ss_intention = 
      across(starts_with("it"),~as.numeric(.)) %>%  
      rowMeans(na.rm = TRUE) %>% 
      scale() %>% as.numeric(),
    #ss_donation as only measuring bd02
    ss_donation = 
      across(all_of(c("bd02")), ~as.numeric(.)) %>%
      rowSums(na.rm = TRUE)
    )

# To make bootstrapping run there cannot be any unordered factors
data.fit2 <- data.fit %>% 
  mutate(
    de01 = as.ordered(de01)
  ) %>% # and neither any factors at all
mutate(
  across(everything(),as.numeric)
)  # but this also good for plotting correlations

summary(data.fit)

#afterdata adding in additional exclusion criteria for bd02 & bd03
#afterdata adding in standardisation of bd02, bdo3 and bd04 prior to scale creation
#afterdata creating sum score variable of donations (bd02 + bd03 + bd04) to be used in analysis as per preregistration
```

```{r}
corrplot(cor(data.fit2,use="pairwise.complete.obs"))
```

## Missing data analysis

Check proportion of missing in all variables

```{r eval=FALSE}

gg_miss_var(
  data.fit2
  )
```

Check major combinations of missingmess.

```{r eval=FALSE}

gg_miss_upset(data.fit2,nsets = 10,nintersects=40)
```

Check if missingness is related to data accumulation.

```{r eval=FALSE}

gg_miss_case_cumsum(
  data.fit2 %>% arrange(dttma)
)

```

Check if missingness is related to major endpoint of the study i.e. intention to donate blood.

```{r eval=FALSE}

gg_miss_fct(
  x = data.fit %>% select(-dttma,
                          -de01,
                          -de02,
                          -bd02,
                          -bd04,
                          -ss_psycown,
                          -ss_selfid,
                          -ss_donation,
                          ), 
  fct = ss_intention)

```

Carry out a crude overall test of missingness.

```{r eval=FALSE}

data.fit3 <- data.fit2 %>%
  select(-ss_psycown,
         -ss_selfid, 
         -ss_intention,
         -ss_donation)

corrplot(cor(data.fit3,use="pairwise.complete.obs"))
mcar_test(data.fit3 %>% select(-dttma))

#afterdata mcar not running, checking if items are too highly correlated
    #taking out sum scores as these were highly correlated with items that made them (i.e., ss_psyown highly correlated with po01, po02, po03)

#afterdata curious what mcar looks like without eO1
#data.fit4 <- data.fit2 %>%
 # select(-ie01)
#mcar_test(data.fit4 %>% select(-dttma))

#mcar_test(data.fit2 %>% select(-dttma)) #doesn't run, likely due to high correlations
# p < 0.05 -> Reject the null hypothesis-> Data is MAR (Missing at random) 
# or MNAR (Missing not at random), meaning missingness follows a pattern.
# But this assumes normally distributed data and 
# does not prove that data is MCAR (Missing Completely at Random)

#afterdata removing sum score donation variable (ss_donation)
```

## Confirmatory factor analysis

```{r}

data.fit2 <- data.fit2 %>%
  mutate(
    ai01 = as.numeric(as.character(ai01)),  # convert to numeric if factor
    #ai01 = ai01 - 1,                         # subtract 1 so lowest category is now 1
    #ai01 = factor(ai01, ordered = TRUE)      # convert back to ordered factor
  )
table(data.fit$ai01)

#afterdata model comparison won't run because no participants responded with "1" for ai01 ("I am intimately familiar with what is going on with regard to donating blood")
  #I'm not sure why but if you just specify that ai01 is numeric it runs without needed to merge level 1 and 2. 
```

```{r}

#making bd02 the only variable in donation history

po01.model <- ' po =~ po01 + po02 + po03
                si =~ si01 + si02 + si03
                it =~ it01 + it02 + it03
                ac =~ ac01 + ac02 + ac03 + ac04
                ai =~ ai01 + ai02 + ai03 + ai04 
                iv =~ iv01 + iv02 + iv03
                bd =~ bd02

'

fit.po01.cfa <- cfa(po01.model,  
                    estimator = "MLR", #
                     missing = "listwise",
                    data = data.fit2)
fit.po01.cfa
```

```{r}
lavaan::summary(fit.po01.cfa,standardized = TRUE) # tidySEM is loaded this fails!
```

### CFA metrics

```{r}
cfastd <- lavaan::standardizedsolution(fit.po01.cfa)
cfastd %>% 
  #"est.std" (or Std.all in above summary output) is "factor loading" and crude rule if thumb is that abs(Std.all) should be above 0.3
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op == '=~')
```

```{r}
lavaan::fitmeasures(fit.po01.cfa,c("npar","chisq","df","cfi","rmsea","srmr"))
```

https://stats.stackexchange.com/questions/541707/cfa-in-lavaan-wont-converge 
"For the comparative fit indices CFI and TLI, you need at least >.900. For the absolute fit indices RMSEA and SRMR, you should have <.080."
So ones from simulated data look good enough.

```{r , fig.height=20}
lavaanPlot(
  fit.po01.cfa,
   edge_options = list(color = "grey"),
  coefs = TRUE,
  graph_options = list(rankdir = "LR")
           ) %>% lavaanPlot::embed_plot_pdf(here("figures/cfa01.pdf"))
```

```{r}
#Make a model with out prespecified correlation structure
fit.po01.cfa.o <- cfa(po01.model,data.fit2,orthogonal=TRUE)
round(
  sapply(list(cfa=fit.po01.cfa,cfa.o=fit.po01.cfa.o),
         function(x) fitmeasures(x,c("npar","chisq","df","pvalue","cfi","rmsea","srmr"))
         )
,3)

#cfa should be better than cfa.o in all measures. 
#cfi and chisq/df should be bigger, rmsea and srmr smaller in model cfa
```

```{r}
#knitr::knit_exit()
#Stop here with real data and review results before modelling.
```

## Sum score SEM

```{r}

include=FALSE

#https://stats.stackexchange.com/questions/340857/serial-mediation-in-r-how-to-setup-the-model

# N.B. you need to check that first item is positively correlated with your latent factor
# Otherwise you get opposite direction

po02ss.model <- ' 
#mediation
ss_psycown ~ a1 *  ss_donation
ss_selfid ~ a2 * ss_donation + d21 * ss_psycown

#regression
ss_intention ~  de01 + de02 + ie01 +  cp * ss_donation + b1 * ss_psycown + b2 * ss_selfid

ind_eff := a1* d21 * b2 
'

cat(po02ss.model)

#afterdata using sum score donation behaviour variable as per preregistered hypothesis

```

```{r echo = FALSE}
#Bivariate correlation table for paper

```

# a) Sum score model with robust estimator and listwise deletion

```{r echo = FALSE}

#if (fitmodels) {
  #fit.po02ss <- sem(po02ss.model,  
                    #estimator = "MLR", 
                    #missing = "listwise",
                    #data = data.fit2,
                    #auto.fix.first = FALSE, 
                    #std.lv = TRUE,
                    #fixed.x = TRUE

 # )

  #models[["fit.po02ss"]] <- fit.po02ss
#}

#lavaan::summary(fit.po02ss,standardized = TRUE)

```

```{r echo = FALSE}

#lavaan::inspect(fit.po02ss,"r2")

```

```{r echo = FALSE}

#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
#fit.po02ss.est %>% select(-label,se,z) %>%  filter(op == '~')

```

```{r echo = FALSE}

#fit.po02ss.est  <-standardizedSolution(fit.po02ss)
#fit.po02ss.est %>% 
  #filter(abs(est.std) < 0.3) %>% 
  #select(-z,-se) %>% 
  #arrange(abs(est.std)) %>% 
  # Just the loadings
  #filter(op != '~1')

```

```{r}

labelsforplottingnull <- c(
  "ss_psycown" = "Psychological ownership",
  "ss_selfid"  = "Self-identity",
  "ss_intention" = "Intention",
  "ss_donation" = "Blood donation history"
)

#lavaanPlot(
  #model = fit.po02ss, 
   #labels = labelsforplottingnull,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem02.pdf"))
```

###  b) Sum score model with bootstrapping and listwise deletion

```{r}
nBoots <- 1000

if (fitmodels) {
  fit.po03ss <- sem(po02ss.model,  
                    estimator = "ML",  
                    missing = "listwise",
                    data = data.fit2, 
                    bootstrap = nBoots,
                    se = "boot",
                    auto.fix.first = FALSE,
                    std.lv = TRUE,
                    fixed.x = TRUE
  )
  models[['fit.po03ss']] <- fit.po03ss
}

lavaan::summary(fit.po03ss,standardized = TRUE)

```

```{r}
fit.po03ss.best  <-standardizedSolution(fit.po03ss)
fit.po03ss.best %>% select(-label,se,z) %>%  filter(op == '~')
```

```{r}
fit.po03ss.best %>% 
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op != '~1')

```

```{r}
lavaanPlot(
  model = fit.po03ss, 
   labels = labelsforplottingnull,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem03.pdf"))

```

#  c) Sum score model with robust estimator and fiml

```{r echo = FALSE}
#if (fitmodels) {
  #fit.po04ss <- sem(
    #po02ss.model,  
    #estimator = "MLR", 
    #data = data.fit2, 
    #auto.fix.first = FALSE,
    #std.lv = TRUE,
    #missing = "fiml",
    #fixed.x = FALSE 
  #)
  #models[['fit.po04ss']] <- fit.po04ss
#}

#lavaan::summary(fit.po04ss,standardized = TRUE)

```

```{r echo = FALSE}

#fit.po04ss.est  <-standardizedSolution(fit.po04ss)
#fit.po04ss.est %>% select(-label,se,z) %>%  filter(op == '~')

```

```{r echo = FALSE}

#fit.po04ss.est %>% 
  #filter(abs(est.std) < 0.3) %>% 
  #select(-z,-se) %>% 
  #arrange(abs(est.std)) %>% 
  # Just the loadings
  #filter(op != '~1')

```

```{r echo = FALSE}

#lavaanPlot(
  #model = fit.po04ss, 
   #labels = labelsforplottingnull,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem04.pdf"))

```

#  d) Sum score model with bootstrapping and fiml

```{r echo = FALSE}

#if (fitmodels) {
  #fit.po05ss <- sem(
    #po02ss.model,  
    #estimator = "ML", 
    #bootstrap = nBoots,
    #se = "boot",
    #data = data.fit2, 
    #auto.fix.first = FALSE,
    #std.lv = TRUE,
    #missing = "fiml",
    #fixed.x = FALSE 
  #)
  #models[['fit.po05ss']] <- fit.po05ss
#}

#lavaan::summary(fit.po05ss,standardized = TRUE)

```

```{r echo = FALSE}

#fit.po05ss.est  <-standardizedSolution(fit.po05ss)
#fit.po05ss.est %>% select(-label,se,z) %>%  filter(op == '~')

```

```{r echo = FALSE}

#fit.po05ss.est %>% 
  #filter(abs(est.std) < 0.3) %>% 
  #select(-z,-se) %>% 
  #arrange(abs(est.std)) %>% 
  # Just the loadings
  #filter(op != '~1')

```

```{r echo = FALSE}

#lavaanPlot(
  #model = fit.po05ss, 
   #labels = labelsforplottingnull,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/plsssem05.pdf"))

```

## Latent factor SEM

```{r}

#making bd02 the only variable in donation history

po02.model <- ' 
#measurement model
psycown =~ po01 + po02 + po03
selfid =~ si01 + si02 + si03
intention =~ it01 + it02 + it03
blooddonor =~ bd02

#bd01 is just used for exclusion of "no"

#mediation
psycown ~ a1 *  blooddonor
selfid ~ a2 * blooddonor + d21 * psycown

#regression
intention ~  de01 + de02 + ie01 +  cp * blooddonor + b1 * psycown + b2 * selfid
ind_eff := a1* d21 * b2
'

cat(po02.model)

```

# e) Latent factor model with robust estimator and listwise deletion

```{r echo = FALSE}
#if (fitmodels) {
  #fit.po02 <- sem(po02.model,  
                  #estimator = "MLR",
                  #missing = "listwise",
                  #data = data.fit2,
                  #auto.fix.first = FALSE,
                  #std.lv = TRUE,
                  #fixed.x = TRUE
  #)
  #models[["fit.po02"]] <- fit.po02
#}
#lavaan::summary(fit.po02,standardized = TRUE)

```

```{r echo = FALSE}

#lavaan::inspect(fit.po02,"r2")

```

```{r echo = FALSE}

#fit.po02.est  <-standardizedSolution(fit.po02)
#fit.po02.est %>% select(-label,se,z) %>%  filter(op == '~')

```

```{r echo = FALSE}

#fit.po02.est %>% 
  #filter(abs(est.std) < 0.3) %>% 
  #select(-z,-se) %>% 
  #arrange(abs(est.std)) %>% 
  # Just the loadings
  #filter(op != '~1')

```

```{r echo = FALSE}
labelsforplotting <- c(
  "psycown" = "Psychological ownership",
  "selfid"  = "Self-identity",
  "intention" = "Intention",
  "blooddonor" = "Blood donation history"
)

#lavaanPlot(
  #model = fit.po02, 
   #labels = labelsforplotting,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem02.pdf"))

```

# f) Latent factor model with boostrapping and listwise deletion

```{r echo = FALSE}

#if (fitmodels) {
  #fit.po03 <- sem(po02.model,  
                  #estimator = "ML", 
                  #data = data.fit2,
                  #bootstrap = nBoots,
                  #se = "boot",
                  #missing = "listwise",
                  #auto.fix.first = FALSE,
                  #std.lv = TRUE,
                  #fixed.x =TRUE
  #)
 # models[['fit.po03']] <- fit.po03
#}

#lavaan::summary(fit.po03,standardized = TRUE)

```

```{r echo = FALSE}
#fit.po03.best <- parameterEstimates(fit.po03)
#fit.po03.best %>% select(-se,-z) %>% filter(op == "~")

```

```{r echo = FALSE}
#lavaanPlot(
  #model = fit.po03, 
   #labels = labelsforplotting,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem03.pdf"))

```

###  g) Latent factor model with robust estimator and fiml

```{r}

if (fitmodels) {
  fit.po04 <- sem(
    po02.model,  
    estimator = "MLR",  
    data = data.fit2,
    auto.fix.first = FALSE,
    std.lv = TRUE,
    missing = "fiml",
    fixed.x = FALSE 
  )
  models[['fit.po04']] <- fit.po04
}

lavaan::summary(fit.po04,standardized = TRUE)

```

```{r}
fit.po04.est  <-standardizedSolution(fit.po04)
fit.po04.est %>% select(-label,se,z) %>%  filter(op == '~')
```

```{r}
fit.po04.est %>% 
  filter(abs(est.std) < 0.3) %>% 
  select(-z,-se) %>% 
  arrange(abs(est.std)) %>% 
  # Just the loadings
  filter(op != '~1')

```

```{r}
lavaanPlot(
  model = fit.po04, 
   labels = labelsforplotting,
  node_options = list(shape = "box", fontname = "Helvetica"), 
  edge_options = list(color = "grey"), 
  coefs = TRUE, 
  stand = TRUE,
  sig=0.001
  ) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem04.pdf"))

```

### Exploratory Analysis (a1*b1 comparison)

```{r}

#factor_scores <- lavPredict(fit.po04, type = "lv") #taking factor scores from CFA 
factor_scores <- lavPredict(fit.po01.cfa, type = "lv")

dim(factor_scores)
dim(data.fit2)

factor_scores <- lavPredict(fit.po04, type = "lv")
head(factor_scores)
summary(factor_scores)

df_scores <- cbind(
  as.data.frame(factor_scores),
  de02 = data.fit2$de02, #de02 = age
  de01 = data.fit2$de01 #de01 = gender
)

df_scores$de01 <- factor(df_scores$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)

#psycown ~ blooddonor
factor1 <- ggplot(df_scores, aes(x = psycown, y = blooddonor)) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor scores: psycown ~ blooddonor",
       x = "Ownership",
       y = "Donation history",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

#selfid ~ blooddonor
factor2 <- ggplot(df_scores, aes(x = selfid, y = blooddonor)) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor scores: selfid ~ blooddonor",
       x = "Identity",
       y = "Donation history",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

#selfid ~ psycown
factor3 <- ggplot(df_scores, aes(x = psycown, y = selfid)) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor scores: selfid ~ psycown",
       x = "Ownership",
       y = "Identity",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

#intention ~ blooddonor
factor4 <- ggplot(df_scores, aes(x = intention, y = blooddonor,
                      color = as.factor(de01))) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor scores: intention ~ blooddonor",
       x = "Intention",
       y = "Donation History",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

#intention ~ psycown
factor5 <- ggplot(df_scores, aes(x = psycown, y = intention,
                      color = as.factor(de01))) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor Scores: psycown ~ intention",
       x = "Ownership",
       y = "Intention",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

#intention ~ selfid
factor6 <- ggplot(df_scores, aes(x = selfid, y = intention,
                      color = as.factor(de01))) +
  geom_point(aes(color = de02), alpha = 0.7, size = 2.5) +
  geom_smooth(aes(linetype = de01), colour = "black", se = FALSE) +
  labs(title = "Factor Scores: selfid ~ intention",
       x = "Identity",
       y = "Intention",
       color = "Age",
       linetype = "Gender") +
  theme_minimal()

combined <- grid.arrange(factor1, factor2, factor3, factor4, factor5, factor6, ncol = 2)

ggsave(
  "Ownership and Intention Factor Loadings",
  combined, width = 20, 
  height = 10, 
  device = "pdf")

```

### exploratory analysis (multigroup analysis to explore impact of gender)

```{r}

#making gender a factor and ordering

data.fit2$de01 <- factor(data.fit2$de01, levels = c(1, 2), labels = c("man", "woman"), ordered = TRUE)

data.fit2 %>%
  group_by(de01) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE)))

#defining new model that doesn't contained de01

ea2.model <- ' 
#measurement model
psycown =~ po01 + po02 + po03
selfid =~ si01 + si02 + si03
intention =~ it01 + it02 + it03
blooddonor =~ bd02

# regressions (separate labels per group)
psycown ~ c(a1g1, a1g2) * blooddonor
selfid ~ c(a2g1, a2g2) * blooddonor + c(d21g1, d21g2) * psycown
intention ~ c(cpg1, cpg2) * blooddonor + c(b1g1, b1g2) * psycown + c(b2g1, b2g2) * selfid

#g1 indirect effects
ind_eff_g1 := a1g1 * d21g1 * b2g1
ind_eff1_g1:= a1g1 * b1g1
ind_eff2_g1:= a2g1 * b2g1

#g2 indirect effects
ind_eff_g2:= a1g2 * d21g2 * b2g2
ind_eff1_g2:= a1g2 * b1g2
ind_eff2_g2:= a2g2 * b2g2
'

ea2.free <- sem(
  ea2.model,
  data = data.fit2,
  group = "de01",
  estimator = "MLR",
  std.lv = TRUE,
  missing = "fiml",
  fixed.x = FALSE,
  auto.fix.first = FALSE,
  group.equal = c('loadings','intercepts')
  )

summary(ea2.free)

ea2.fixed <- sem(
  ea2.model,
  data = data.fit2,
  group = "de01",
  estimator = "MLR",
  std.lv = TRUE,
  missing = "fiml",
  fixed.x = FALSE,
  auto.fix.first = FALSE,
  group.equal = c('loadings', 'intercepts', 'regressions')
)

# this is for model comparison

lavTestLRT(ea2.free, ea2.fixed)

#ea2 <- sem(
  #ea2.model,
  #data = data.fit2,
  #group = "de01",
  #estimator = "MLR",
  #std.lv = TRUE,
  #missing = "fiml",
  #fixed.x = FALSE,
  #auto.fix.first = FALSE
#)

#summary(ea2, standardized = TRUE)

#end of exploratory analysis
knitr::knit_exit()

```

###  h) Latent factor model with bootstrapping and fiml

```{r echo = FALSE}

#if (fitmodels) {
  #fit.po05 <- sem(
    #po02.model,  
    #estimator = "ML",  
    #data = data.fit2, 
    #bootstrap = nBoots,
    #se = "boot",
    #auto.fix.first = FALSE,
    #std.lv = TRUE,
    #missing = "fiml",
    #fixed.x = FALSE 
  #)
  #models[['fit.po05']] <- fit.po05
#}

#lavaan::summary(fit.po05,standardized = TRUE)

```

```{r echo = FALSE}
#fit.po05.est  <-standardizedSolution(fit.po05)
#fit.po05.est %>% select(-label,se,z) %>%  filter(op == '~')

```

```{r echo = FALSE}
#fit.po05.est %>% 
  #filter(abs(est.std) < 0.3) %>% 
  #select(-z,-se) %>% 
  #arrange(abs(est.std)) %>% 
  # Just the loadings
  #filter(op != '~1')

```

```{r echo = FALSE}
#lavaanPlot(
  #model = fit.po05, 
   #labels = labelsforplotting,
  #node_options = list(shape = "box", fontname = "Helvetica"), 
  #edge_options = list(color = "grey"), 
  #coefs = TRUE, 
  #stand = TRUE,
  #sig=0.001
  #) %>% lavaanPlot::embed_plot_pdf(here("figures/pllvsem05.pdf"))

```

## Compare models

### Effect sizes

```{r , fig.height=8}

data.forest  <- fit.po03.best %>% 
  mutate(
    estmet = "Bootstrapped",
    missing = "Listwise"
         ) %>%  
  #rename(est=est.std) %>% 
  filter(op == "~" | op == ":=") %>% 
  bind_rows(
    fit.po02.est %>% mutate(
      estmet = "Robust",
      missing = "Listwise"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":="),
    fit.po04.est %>% mutate(
      estmet = "Robust",
      missing = "FIML"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":="),
    fit.po05.est %>% mutate(
      estmet = "Boostrapped",
      missing = "FIML"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=")

  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs),
    linetype = "Latent factor"
    ) %>% 
  arrange(estmet, term)

data.forest  <- fit.po03ss.best %>% 
  rename(est=est.std) %>% 
  mutate(
    estmet = "Bootstrapped",
    missing = "Listwise"
    ) %>%  
  filter(op == "~"| op == ":=") %>% 
  bind_rows(
    fit.po02ss.est %>% mutate(
      estmet = "Robust",
      missing = "Listwise"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=") ,
    fit.po04ss.est %>% mutate(
      estmet = "Robust",
      missing = "FIML"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":="),
    fit.po05.est %>% mutate(
      estmet = "Boostrapped",
      missing = "FIML"
      ) %>% 
      rename(est=est.std) %>% 
      filter(op == "~"| op == ":=")

  ) %>% 
  mutate(
    term = str_c(label,":",lhs,op,rhs),
    linetype = "Sum score"
    ) %>% 
  arrange(estmet, term) %>% 
  bind_rows(data.forest) %>% 
  mutate(
    lhs =str_replace_all(lhs,"ss_",""),
    rhs =str_replace_all(rhs,"ss_",""),
    term =str_replace_all(term,"ss_",""),
    term =str_replace_all(term,"ind_eff:ind_eff:","ind_eff:"),
    term =str_replace_all(term,"ss_donation","blooddonor:"),
    term =str_replace_all(term,"blooddonor:","blooddonor"),
    term =str_replace_all(term,"^:",""),
    estmet = as.factor(estmet),
    mtype = as.factor(str_c(estmet,"-",linetype, "-",missing)),
    color = as.factor(str_c(estmet," and ",missing)),
    hollow_group=factor(if_else(ci.lower<=0 & ci.upper>=0, NA_character_, as.character(color)),
                             levels=levels(color))
  )

pos <- position_nudge(y=as.numeric(as.factor(data.forest$mtype))/10 -0.5)

pf <- data.forest %>% 
  ggplot(aes(est,xmin=ci.lower,xmax=ci.upper,y=term,fill=hollow_group,color=color,linetype = linetype))+
  geom_vline(xintercept=0,color="grey") +
  geom_linerange(position=pos)+
  geom_point(shape=21,size=3,position=pos,alpha=0.7)+
  scale_fill_discrete(na.value=NA, guide ="none") + 
  labs(x="Standardised effect size",y="Relationship") +
  labs(color="Model type", linetype="Variable type") +
  geom_stripped_rows(color=NA) + theme_classic()

pf
```

```{r}
ggsave(pf,
       filename = here("figures/forest01.png"),
       height = 250,
       width = 180,
       units = "mm"
         )

ggsave(pf,
       filename = here("figures/forest01.pdf"),
       height = 250,
       width = 180,
       units = "mm"
         )

```

## Indices of goodness of fit

```{r}
if (fitmodels) {
  gfi<- round(
    sapply(
      list(
        SSRL=fit.po02ss,
        SSBL=fit.po03ss,
        SSRF = fit.po04ss,
        SSBL = fit.po05ss,
        LFRL=fit.po02,
        LFBL=fit.po03, 
        LFRF=fit.po04,
        LFBF=fit.po05
      ),
      function(x) fitmeasures(x,c("chisq","df","pvalue","cfi","rmsea","srmr"))
    )
    ,3)
  models[['gfi']] <- gfi
} else {
  gfi <- models$gfi
}
gfi
```

According to https://easystats.github.io/effectsize/reference/interpret_gfi.html

"For structural equation models (SEM), Kline (2015) suggests that at a minimum the following indices should be reported: The model chi-square, the RMSEA, the CFI and the SRMR."

chisq, Chi-squared, chisq/df ratio smaller is better

cfi Comparative Fit Index, "It should be > .96" , higher is better

rmsea Root Mean Square Error of Approximation, "should be < .08", smaller is better

srmr Standardized Root Mean Square Residual, "should be < .08", smaller is better

NB. Sum score and latent factor models use different data and are not nested hence they cannot be compared with these metrics.

```{r}
if (fitmodels) {
  save(models,file=here("prereg_models.rdata"))

}
```